### 2019-09-05 manuscript/manuscript.tex
G's changes

Finally, we observed that in the longest run evolve-and-resequence study we
analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
appeared to become negative at future timepoints (see the first two rows in
Figure \ref{fig:figure-1} A). This could be caused by advantageous fitness
backgrounds later becoming disadvantageous, either due to frequency-dependent
selection, recombination breaking up coupled beneficial and deleterious
alleles, or an environmental change in the direction of selection. However, the
signal of a change in the direction of selection localized to a particular
region of the genome may be washed out when we calculate genome-wide temporal
covariances.  Averaging over such large numbers of loci is necessary to detect
a signal of covariance in allele frequency changes when both drift (inflated in
evolve-and-sequence studies which use small population sizes) and sampling
variance can create spurious covariances. To address this limitation, we
calculated temporal covariances over 100kb genomic windows (mean loci per
window 4,606, median 4,830). While the covariance of each tile is noisy, we can
analyze the genome-wide distribution of tile covariances for changes through
time. We would like to compare the distribution of these tile covariances to a
neutral expectation, but using a theoretic neutral null distribution would be
inappropriate given complex dependencies across the genome due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).

Instead, we have developed a permutation-based procedure that constructs a null
distribution of the genome-wide covariances as they would look if only neutral
genetic drift was acting. We do this by randomly flipping the signs of the
allele frequency changes per-genomic window 1,000 times, which acts to destroy
the systematic covariances created by linked selection but creates a sampling
distribution of the covariances spuriously created by neutral genetic drift.
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift alone, since
drift also acts to inflate the magnitude of allele frequency change. By
flipping the sign at the block-level, we preserve the complex dependencies
between adjacent loci created by linkage disequilibrium; see Supplementary
Material Section \ref{supp:empirical-null} for more details on this approach.

Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
null to investigate whether there was a general trend for the temporal
covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
expected under selection), but then become systematically negative for larger
$k$ (as would occur under evolutionary processes that reverse the fitness of
regions across the genome). Such a change in the fitness direction of
haplotypes would not be discernible from genome-wide covariances like those in
Figure \ref{fig:figure-1}. We find (Figure \ref{fig:figure-3} A and B), pooling
across all replicates (see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates), that
windowed temporal covariances between close timepoints are skewed positive (a
heavy right tail), while between more distant timepoints these windowed
temporal covariances tend to shift to become more negative (a heavy left tail).
We additionally quantified the degree to which the left and right tails change
as we look at covariances between timepoints further apart (Figure
\ref{fig:figure-3} C). We find that for covariances 2 timepoints apart ($k=2
\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
[1.29, 1.32]), consistent with linked selection creating positive covariance in
allele frequency change. Under constant directional selection, we expect these
positive covariances between increasingly distant timepoints to decay to zero
as additive genetic variance is exhausted by selection and linkage
disequilibrium decays. Instead we find that for covariances 4 timepoints apart
($k=4 \times 10$ generations), the left tail of negative covariances is
inflated by a factor of 1.19 (95\% CI [1.17, 1.21]). This finding is also
robust to sign-permuting allele frequency changes on a chromosome-level, the
longest extent that gametic linkage disequilibria can extend (Supplementary
Figure \ref{suppfig:barghi-tailprobs-seqid}). Overall, we see a striking
pattern that covariances not only decay towards zero, but in fact become
negative through time, consistent with some regions in the genome having had a
reversed fitness effect at later timepoints.

@@

  \caption{A, B: The distribution of temporal covariances calculated in 100kb
    genomic windows from the \textcite{Barghi2019-qy} study, plotted alongside
    an empirical neutral null distribution created by recalculating the
    windowed covariances on sign permuting the allele frequency changes. The
    bin number is 88, chosen by cross validation (Supplementary Materials
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure A, windowed
    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
    10$ generations and in subfigure B the covariances are separated by $k=4
    \times 10$ generations; each $k$ is an off-diagonal from the variance
    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
    of covariance matrix in subfigures A and B, where the first diagonal is the
    variance, and the dark gray indicates which off-diagonal of the covariance
    matrix is plotted in the histograms). C: The lower and upper tail
    probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
light gray dashed line indicates the 20\% tail probability expected under the
neutral null.}

@@

>>>>

Finally, we observed that in the \gc{longest study}  we   %run evolve-and-resequence 
analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
\gc{become} negative at future timepoints (see the first two rows in
Figure \ref{fig:figure-1} A). \gc{This shows that alleles that were on
average going up initially are no going down in frequency, i.e. that
the average direction of selection experienced by alleles has
flipped. This must reflect either a change in the environment or the
genetic background, due to changing epistatic relationships among
alleles or  recombination breaking up selective alleles.}
\gc{Such reversals of selective dynamics could be occuring at other time-points but} the
signal of a change in the direction of selection \gc{at particular
  loci} may be washed out when we calculate our  genome-wide \gc{average} temporal
covariances. To address this limitation, we
calculated \gc{the distribution of the} temporal covariances over
100kb genomic windows \gc{(Figure \ref{fig:figure-3}, pooling
across all replicates see Supplementary Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).}
\gc{The}
covariance of each tile is noisy, \gc{due to sampling and genetic
  drift, shown in }. \gc{The neutral distribution of the covariance is
  complicated due to} linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). 
\gc{Thus} we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the    %of the
                                %genome-wide covariances as they would
                                %look if only neutral genetic drift was acting. We do this 
allele frequency changes per-genomic window. \gc{This destroys}  %%  GC bit about 1,000 times seems more like figgure caption or methods
the systematic covariances created by linked selection but creates a sampling
distribution of the covariances spuriously created by neutral genetic
drift while preserve the complex dependencies
between adjacent loci created by linkage disequilibrium (see Supplementary
Material Section \ref{supp:empirical-null} for more details on this approach).
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift
alone \gc{as they include the effect of selection on the alele
  frequency change within a time-interval, just not between
  time-intervals. We see, Figure \ref{fig:figure-3} A and B, that
windowed temporal covariances between close timepoints are skewed positive (a
heavy right tail), while between more distant timepoints these windowed
temporal covariances tend to shift to become more negative (a heavy
left tail).
We quantified the degree to which the left and right
tails \gc{are inflated compared to the null distribution as a function
  of time}, and see excesses in both tails in Figure
\ref{fig:figure-3} C. This finding is also
robust to sign-permuting allele frequency changes on a chromosome-level, the
longest extent that gametic linkage disequilibria can extend {\bf ????} (Supplementary
Figure \ref{suppfig:barghi-tailprobs-seqid}).  \gc{We} see a striking
pattern that \gc{the windowed} covariances not only decay towards zero, but in fact become
negative through time, consistent with some \gc{many} regions in the genome having had a
reversed fitness effect at later timepoints.

%We find that for covariances 2 timepoints apart ($k=2
%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
%[1.29, 1.32]), consistent with linked selection creating positive covariance in
%allele frequency change. Under constant directional selection, we expect these
%positive covariances between increasingly distant timepoints to decay to zero
%as additive genetic variance is exhausted by selection and linkage
%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
%($k=4 \times 10$ generations), the left tail of negative covariances is
%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).

@@

  \caption{A, B: The distribution of temporal covariances calculated in 100kb
    genomic windows from the \textcite{Barghi2019-qy} study, plotted alongside
    an empirical neutral null distribution created by recalculating the
    windowed covariances on \gc{a 1000 sign permutions} allele frequency changes. The
    bin number is 88, chosen by cross validation (Supplementary Materials
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure A, windowed
    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
    10$ generations and in subfigure B the covariances are separated by $k=4
    \times 10$ generations; each $k$ is an off-diagonal from the variance
    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
    of covariance matrix in subfigures A and B, where the first diagonal is the
    variance, and the dark gray indicates which off-diagonal of the covariance
    matrix is plotted in the histograms). C: The lower and upper tail
    probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
light gray dashed line indicates the 20\% tail probability expected under the
neutral null.}

@@

%since
%drift also acts to inflate the magnitude of allele frequency change. By
f%lipping the sign at the block-level, we

%we can  % (mean \gc{SNP} perwindow 4,606, median 4,830)
%analyze the genome-wide distribution of tile covariances for changes through
t%ime. We would like to compare the distribution of these tile covariances to a
%neutral expectation, but using a theoretic neutral null distribution would be
%inappropriate given complex dependencies across the genome due to linkage
%disequilibria (which can occur over long physical distances in E\&R and
%selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).

%Averaging over such large numbers of loci is necessary to detect
%a signal of covariance in allele frequency changes when both drift (inflated in
%evolve-and-sequence studies which use small population sizes) and sampling
%variance can create spurious covariances.

%This could be caused by advantageous fitness
%backgrounds later becoming disadvantageous, either due to frequency-dependent
%selection, recombination breaking up coupled beneficial and deleterious
%alleles, or an environmental change in the direction of selection.
%However, the
%signal of a change in the direction of selection localized to a particular
%region of the genome may be washed out when we calculate genome-wide temporal
%covariances.

%Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
%null to investigate whether there was a general trend for the temporal
%covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
%expected under selection), but then become systematically negative for larger
%$k$ (as would occur under evolutionary processes that reverse the fitness of
%regions across the genome). Such a change in the fitness direction of
%haplotypes would not be discernible from genome-wide covariances like those in
%Figure \ref{fig:figure-1}.
